---
title: "Day3_furrr"
author: "bahapku"
date: "2020/4/4"
output: 
    prettydoc::html_pretty:
      theme: hpstr
      highlight: github
---
#different data ,one model
```{r}
library(broom)
library(furrr)
library(tidyr)
library(gapminder)
library(tictoc)
plan(multiprocess)
tic()
gapminder %>% group_by(country) %>% nest() %>%ungroup() %>% 
  mutate(models=future_map(data,~lm(lifeExp~pop+gdpPercap,data = .x)),
         detail_coefs=models %>% future_map(tidy,.options = future_options(packages = "broom")),
         simple_coefs=models %>% future_map(coef),
         intercept=models %>% future_map_dbl(~pluck(coef(.),"(Intercept)")),
         coef_yours=models %>% future_map_dbl(~pluck(coef(.),"pop")),
         stat_inf=models %>% future_map(glance,.options = future_options(packages = "broom")),
         augments=models %>% future_map(augment,.options = future_options(packages = "broom")),
         AIC=stat_inf %>% future_map_dbl("AIC"),
         BIC=stat_inf %>% future_map_dbl("BIC"),
         degree_fredom=stat_inf %>% future_map_dbl("df"),
         r_squared=stat_inf %>% future_map_dbl("r.squared"),
         P_value=stat_inf %>% future_map_dbl("p.value"))
          
  toc()

  
```
仅仅运用了6秒时间    

```{r}
library(broom)
library(furrr)
library(tidyr)
library(gapminder)
library(tictoc)
plan(multiprocess)
gapminder %>% group_by(country) %>% nest() %>%ungroup() %>% 
  mutate(models=future_map(data,~lm(lifeExp~pop+gdpPercap,data = .x)),
         detail_coefs=models %>% future_map(tidy,.options = future_options(packages = "broom")),
         simple_coefs=models %>% future_map(coef),
         intercept=models %>% future_map_dbl(~pluck(coef(.),"(Intercept)")),
         coef_yours=models %>% future_map_dbl(~pluck(coef(.),"pop")),
         stat_inf=models %>% future_map(glance,.options = future_options(packages = "broom")),
         augments=models %>% future_map(augment,.options = future_options(packages = "broom")),
         AIC=stat_inf %>% future_map_dbl("AIC"),
         BIC=stat_inf %>% future_map_dbl("BIC"),
         degree_fredom=stat_inf %>% future_map_dbl("df"),
         r_squared=stat_inf %>% future_map_dbl("r.squared"),
         P_value=stat_inf %>% future_map_dbl("p.value")) %>% 
  ggplot()+geom_point(aes(x=country,y=sort(r_squared),color= P_value<0.05))+
  theme_ipsum()
```
也可以紧接着画你想要的任何图    

# one data,different models
```{r}
f1 <- lifeExp ~ pop
f2 <- lifeExp ~ pop+gdpPercap
f3 <- lifeExp ~ pop + gdpPercap + year
f4 <- lifeExp ~ pop + gdpPercap + year + continent
formulas <- list(f1,f2,f3,f4)
library(broom)
library(furrr)
library(tidyr)
library(gapminder)
library(tictoc)
head(gapminder,5)
plan(multiprocess)
tic()
 formulas %>% future_map(~lm(.x,data = gapminder)) -> models
  tibble(
    formulas=models %>% future_map_chr(~format(formula(.))),
    coef_yours=models %>% future_map_dbl(~pluck(coef(.),"pop")),
    r_squared=models %>% future_map_dbl(~pluck(glance(.),"r.squared"),.options = future_options(packages = "broom")),
    p_value=models %>% future_map_dbl(~pluck(glance(.),"p.value"),.options = future_options(packages = "broom"))
  ) 
 toc()
```

#one model ,for each index data
```{r}
library(broom)
library(furrr)
library(tidyr)
library(gapminder)
library(tictoc)
plan(multiprocess)
tic()
gapminder %>% split(.$country)  -> split_data
split_data %>% future_map(~lm(lifeExp ~ pop + gdpPercap + year,data = .x)) %>% 
  future_map_dfr(~tibble(
    formulas=format(formula(.x)),
    detail_coefs=list(tidy(.x)),
    simple_coefs=list(coef(.x)),
    r_squared=pluck(glance(.x),"r.squared"),
    p_value=pluck(glance(.x),"p.value")
    
  ),.id = "country",.options = future_options(packages = "broom"))
toc()
```

# advantages on summerizing     
question: Subset the gapminder data according to the following regions, then find the mean life expectancy and mean GDP per capita for each region in each year    

ANZ <- c("Australia", "New Zealand")   

SEA <- c("Cambodia", "Indonesia", "Malaysia", "Myanmar", "Philippines", 
         "Singapore", "Thailand", "Vietnam")    
         
SA <- c("Bangladesh", "India", "Nepal", "Pakistan", "Sri Lanka")   

EA <- c("China", "Hong Kong, China", "Japan", "Korea, Dem. Rep.", "Korea, Rep.", 
        "Taiwan")   
        
```{r}
library(broom)
library(furrr)
library(tidyr)
library(gapminder)
library(tictoc)
plan(multiprocess)
ANZ <- c("Australia", "New Zealand")
SEA <- c("Cambodia", "Indonesia", "Malaysia", "Myanmar", "Philippines", 
         "Singapore", "Thailand", "Vietnam")
SA <- c("Bangladesh", "India", "Nepal", "Pakistan", "Sri Lanka")
EA <- c("China", "Hong Kong, China", "Japan", "Korea, Dem. Rep.", "Korea, Rep.", 
        "Taiwan")
list(ANZ=ANZ, SEA=SEA, SA=SA, EA=EA) %>% 
 future_map(~dplyr::filter(gapminder,country  %in% .x)) %>% 
  future_map(~group_by(.x,year)) %>% 
  future_map_dfr(~summarise(.x,mean_lifeexp=mean(lifeExp),mean_gdp=mean(gdpPercap)),.id = "region")  -> result
  
  result
ggplot( result, aes(x=year, y=mean_lifeexp, col=region)) +
  geom_line()
  

```

